Perturbing SU($n$)-symmetric interactions

### 2020-01-13 collective_sun/sun_perturbation.tex
revert to previous convention for denoting collective operators, making them extensive

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_0
  \EQFS N \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \f1N \sum_p X_p.
\end{align}
In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on
the diagonal and zeros everywhere else.  Solving the general
eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing the
$N\times N$ matrix $\m h-\diag\v h$.  When the interaction Hamiltonian
$H_0$ is translationally invariant, however, this eigenvalue problem
can be solved analytically; we provide the corresponding solution in
Appendix \ref{sec:trans_inv}.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the fully symmetric manifold
$\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQFS \f{N}{N-1} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-body operator $Z$, we define the spin-averaged
operator
\begin{align}
  \col{Z} \equiv \f1N \sum_p Z_p.
\end{align}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
  \EQFS {N \choose 2} \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv {N \choose 2}^{-1} \sum_{p<q} O_{pq},
\end{align}
are averages of $w_{Opq}$ and $O_{pq}$ over all pairs of spins $p,q$.
If the two-body operator $O=X\otimes Y$, then
\begin{align}
  \col{O} \EQFS X_0 Y_1
  \EQFS \f1{N-1} \p{N\, \col{X}\,\col{Y} - \col{XY}},
\end{align}
so
\begin{align}
  H_2^{(1)} \EQFS \f12 N \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{N\, \col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_{pq} - \f12\p{h_p + h_q},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{pk} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
The eigenvalue problem in \eqref{eq:cond_2} can be simplified somewhat
when the interaction Hamiltonian $H_0$ is translationally invariant;
we discuss this simplification in Appendix \ref{sec:trans_inv}.

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized0 state of
definite energy that depends only on the coefficients $\m w_O^\Delta$.
Similarly to the case of single-body perturbations, we can decompose
any vector $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto
the eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2^\Delta \E \V_2^\Delta \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
then
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to expand
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \p{\V_1^\Delta}^2 \P_0
  \EQFS \sum_{\substack{X,Y\\p,q}} v_{Xp} v_{Yq} X_p Y_q
  = \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
  + \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q,
  \label{eq:PXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; and all coefficient vectors $\v v_X$
satisfy \eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies
that all $\v v_X$ are mean-zero.  The first sum on the right of
\eqref{eq:PXYP_start} is then
\begin{align}
  \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
  \EQFS \sum_{X,Y,p} v_{Xp} v_{Yp} X_0 Y_0
  \EQFS \sum_{X,Y} \v v_X \c\v v_Y \col{XY},
  &&
  \col{Z} \equiv \f1N \sum_p Z_p.
  \label{eq:PXYP_eq}
\end{align}
The second sum on the right of \eqref{eq:PXYP_start}, meanwhile, is
\begin{align}
  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
  \EQFS \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_0 Y_1,
  \label{eq:PXYP_neq_start}
\end{align}
where we can use the fact that the vectors $\v v_X,\v v_Y$ are
mean-zero to simplify the sum over $p\ne q$ in
\eqref{eq:PXYP_neq_start} to
\begin{align}
  \sum_{p\ne q} v_{Xp} v_{Yq}
  = \sum_{p,q} v_{Xp} v_{Yq} - \sum_{p=q} v_{Xp} v_{Yq}
  = - \v v_X\c\v v_Y,
\end{align}
In order to simplify the product of operators in
\eqref{eq:PXYP_neq_start}, we expand
\begin{align}
  N^2 \col{X}\,\col{Y} = \sum_{p,q} X_p Y_q
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQFS N \col{XY} + N\p{N-1} X_0 Y_1,
  \label{eq:PXYP_neq_ops}
\end{align}
which implies that
\begin{align}
  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
  \EQFS - \sum_{X\ne Y} \v v_X\c\v v_Y X_0 Y_1
  \EQFS - \f1{N-1}\sum_{X,Y} \v v_X\c\v v_Y
  \p{N \col{X}\,\col{Y} - \col{XY}}.
  \label{eq:PXYP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \P_0 \p{\V_1^\Delta}^2 \P_0
  = - \f{N}{N-1} \sum_{X,Y} \v v_X \c\v v_Y
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}

@@

spin-average, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}/N$, where $Q^{(j)}$
acts with the single-body operator $Q$ on spin $j$ and trivially on
all other spins.  By construction, the set of collective operators
$\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the set of
multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is
straightforward to work out that for any set $S$ of single-spin
operators,

@@

  N^{\abs{S}} \O_S

@@

  N!_{\abs{S}} O_S

@@

  \p{-1}^{\abs{S}-\abs{P}} N^{\abs{P}} \O_P
  \times \prod_{p\in P}\p{\abs{p}-1}!

@@

  \p{-1}^{\abs{S}-\abs{\tilde P}} N^{\abs{\tilde P}} \O_{\tilde P}

@@

  N^{\abs{\tilde S}} \O_{\tilde S} \EQFS

@@

While the expansions in \eqref{eq:multi_to_collective_full} and
\eqref{eq:multi_to_collective_full_simp} closely resemble each other,
the latter can involve less computational overhead due to the lack of
symmetrization in the construction of $\O_{\tilde P}$.  There are, in
principle, even more economical decompositions of a multi-local
operators into collective opreators, for example
\begin{align}
  N!_2\, X\otimes Y
  &\EQFS N^2\, \col{X}\,\col{Y} - N\, \col{XY}, \\
  N!_3\, X\otimes Y \otimes Z
  &\EQFS N^3\, \col{X}\,\col{Y}\,\col{Z}
  - N^2 \p{\col{X}\,\col{YZ} + \col{Y}\,\col{XZ} + \col{Z}\,\col{XY}}
  + N \p{\col{ZXY} + \col{YXZ}},
\end{align}
but these decompositions generally do not preserve the manifest
symmetry of a multi-local operator $O_S$ under arbitrary permutations
of $S$, and we do not know of any general prescription to arrive at
such decompositions aside from solving systems of equations relating
collective operators to multi-local operators.

On a final note, we consider the special case that all operators
$O\in S$ are identical, for which we provide the expansions
\begin{align}
  N!_2\, X\otimes X \EQFS N^2\, {\col{X}\,}^2 - N\, \col{X^2},
  &&
  N!_3\, X^{\otimes 3} \EQFS N^3\, {\col{X}\,}^3
  - 3 N^2\, \col{X} \odot \col{X^2} + 2 N\, \col{X^3},
\end{align}
\begin{align}
  N!_4\, X^{\otimes 4}
  \EQFS N^4\, {\col{X}\,}^4 - N^3 \p{4 {\col{X}\,}^2 \odot \col{X^2}
    + 2 \col{X} \, {\col{X}\,}^2 \, \col{X}}
  + N^3\p{3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}}
  - 6 N\, \col{X^4}.
\end{align}
If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
then
\begin{align}
  N!_2\, X\otimes X \EQFS N^2\, {\col{X}\,}^2 - N,
  &&
  N!_3\, X^{\otimes 3}
  \EQFS N^3\, {\col{X}\,}^3 - \p{3N-2} N\, \col{X},
\end{align}
\begin{align}
  N!_4\, X^{\otimes 4} \EQFS N^4\, {\col{X}\,}^4
  - 2\p{3N-4} N^2\, {\col{X}\,}^2 + 3 N \p{N-2}.
\end{align}

>>>>

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_0
  \EQFS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on
the diagonal and zeros everywhere else.  Solving the general
eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing the
$N\times N$ matrix $\m h-\diag\v h$.  When the interaction Hamiltonian
$H_0$ is translationally invariant, however, this eigenvalue problem
can be solved analytically; we provide the corresponding solution in
Appendix \ref{sec:trans_inv}.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the fully symmetric manifold
$\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQFS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N\, \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-body operator $Z$, we define the collective operator
\begin{align}
  \col{Z} \equiv \sum_p Z_p.
\end{align}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-body
operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_{pq} - \f12\p{h_p + h_q},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{pk} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
The eigenvalue problem in \eqref{eq:cond_2} can be simplified somewhat
when the interaction Hamiltonian $H_0$ is translationally invariant;
we discuss this simplification in Appendix \ref{sec:trans_inv}.

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized state of
definite energy that depends only on the coefficients $\m w_O^\Delta$.
Similarly to the case of single-body perturbations, we can decompose
any vector $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto
the eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2^\Delta \E \V_2^\Delta \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
then
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to expand
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \p{\V_1^\Delta}^2 \P_0
  \EQFS \sum_{\substack{X,Y\\p,q}} v_{Xp} v_{Yq} X_p Y_q
  = \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
  + \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q,
  \label{eq:PXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; and all coefficient vectors $\v v_X$
satisfy \eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies
that all $\v v_X$ are mean-zero.  The first sum on the right of
\eqref{eq:PXYP_start} is then
\begin{align}
  \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
  \EQFS \sum_{X,Y,p} v_{Xp} v_{Yp} X_0 Y_0
  \EQFS \f1N \sum_{X,Y} \v v_X \c\v v_Y \col{XY},
  &&
  \col{Z} \equiv \sum_p Z_p.
  \label{eq:PXYP_eq}
\end{align}
The second sum on the right of \eqref{eq:PXYP_start}, meanwhile, is
\begin{align}
  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
  \EQFS \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_0 Y_1,
  \label{eq:PXYP_neq_start}
\end{align}
where we can use the fact that the vectors $\v v_X,\v v_Y$ are
mean-zero to simplify the sum over $p\ne q$ in
\eqref{eq:PXYP_neq_start} to
\begin{align}
  \sum_{p\ne q} v_{Xp} v_{Yq}
  = \sum_{p,q} v_{Xp} v_{Yq} - \sum_{p=q} v_{Xp} v_{Yq}
  = - \v v_X\c\v v_Y,
\end{align}
In order to simplify the product of operators in
\eqref{eq:PXYP_neq_start}, we expand
\begin{align}
  \col{X}\,\col{Y} = \sum_{p,q} X_p Y_q
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQFS \col{XY} + N\p{N-1} X_0 Y_1,
  \label{eq:PXYP_neq_ops}
\end{align}
which implies that
\begin{align}
  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
  \EQFS - \sum_{X\ne Y} \v v_X\c\v v_Y X_0 Y_1
  \EQFS - \f1{N\p{N-1}}\sum_{X,Y} \v v_X\c\v v_Y
  \p{\col{X}\,\col{Y} - \col{XY}}.
  \label{eq:PXYP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \P_0 \p{\V_1^\Delta}^2 \P_0
  = - \f1{N\p{N-1}} \sum_{X,Y} \v v_X \c\v v_Y
  \p{\col{X}\,\col{Y} - N\,\col{XY}}.
\end{align}

@@

collective version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where
$Q^{(j)}$ is an operator that acts with the single-spin operator $Q$
on spin $j$ and trivially on all other spins.  By construction, the
set of collective operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is
one-to-one to the set of multi-local operators
$\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is straightforward to
work out that for any set $S$ of single-spin operators,

@@

  \O_S

@@

  N!_{\abs{S}}\, O_S

@@

  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!

@@

  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}

@@

  \O_{\tilde S} \EQFS

@@

While the expansions in \eqref{eq:multi_to_collective_full} and
\eqref{eq:multi_to_collective_full_simp} closely resemble each other,
in practice the latter involves less computational overhead due to the
lack of symmetrization in the construction of $\O_{\tilde P}$.  There
are, in principle, even more economical decompositions of a
multi-local operators into collective opreators, for example
\begin{align}
  N!_2\, X\otimes Y
  &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
  N!_3\, X\otimes Y \otimes Z
  &\EQFS \col{X}\,\col{Y}\,\col{Z}
  - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
  + \col{ZXY} + \col{YXZ},
\end{align}
but these decompositions generally do not preserve the manifest
symmetry of a multi-local operator $O_S$ under arbitrary permutations
of $S$, and we do not know of any general prescription to arrive at
such decompositions aside from solving systems of equations relating
collective operators to multi-local operators.

On a final note, we consider the special case that all operators
$O\in S$ are identical, for which we provide the expansions
\begin{align}
  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - \col{X^2},
  &&
  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3
  - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
\end{align}
\begin{align}
  N!_4\, X^{\otimes 4}
  \EQFS {\col{X}\,}^4 - 4 {\col{X}\,}^2 \odot \col{X^2}
  - 2 \col{X} \, {\col{X}\,}^2 \, \col{X}
  + 3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}
  - 6 \col{X^4}.
\end{align}
If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
then
\begin{align}
  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - N,
  &&
  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3 - \p{3N-2} \col{X},
\end{align}
\begin{align}
  N!_4\, X^{\otimes 4} \EQFS {\col{X}\,}^4
  - 2\p{3N-4} {\col{X}\,}^2 + 3 N \p{N-2}.
\end{align}
